---
title: "PhenoRam analysis"
author: "Cristina Garcia Timmermans & Ruben Props"
date: "Today"
output:
  html_document:
    code_folding: show
    highlight: haddock
    keep_md: yes
    theme: united
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 2
    css: report_styles.css
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, 
                      echo = TRUE, 
                      cache = TRUE,
                      include = TRUE,
                      collapse = FALSE,
                      dependson = NULL,
                      engine = "R", # Chunks will always have R code, unless noted
                      error = TRUE,
                      fig.path="Figures/cached/",  # Set the figure options
                      fig.align = "center",
                      dev.args=list(bg='transparent')
                      )
```

# Load libraries

```{r load-libraries, message=FALSE, warning = FALSE}
library("Phenoflow")
library("mclust")
library("plyr")
library("dplyr")
library("gridExtra")
library("tidyr")
library("phyloseq")
library("ggplot2")
library("gridExtra")
library("fpc")
library("RColorBrewer")
library("vegan")
library("tsne")
library("sandwich")
library("cluster")
library("grid")
library("egg")
library("MALDIquant")
library("reshape2")
library("caret") # for cross validation
library("foreach") # for parallelization
source("functions.R")
library("flowAI") # for denoising of FCM data
load("workspace.RData")

my.settings <- list(
  strip.background=list(col="transparent"),
  strip.border=list(col="transparent", cex=5),
  gate=list(col="black", fill="lightblue", alpha=0.2,border=NA,lwd=2),
  panel.background=list(col="lightgray"),
  background=list(col="white"))
```

# Cluster analysis
## A. Hyperspec-normalized spectra
We will start with the hyperspec processed spectr:  

* Select number of stable clusters in data set
* Map back to original samples
* Calculate phenotypic diversity in each sample using Hill numbers

```{r determine-clusters, fig.width = 7, fig.height= 6, dpi = 500, warning=FALSE}
# Choose if you want to run PCA prior to clustering
PCA <- TRUE

if(PCA == TRUE){
  # Perform PCA to reduce number of features in fingerprint
  pca_bacteria <- prcomp(hs.norm)

  # Only retain PC which explain 90% of the variance
  thresh <- 0.9
  nr_pc_bacteria <- min(which((cumsum(vegan::eigenvals(pca_bacteria)/sum(vegan::eigenvals(pca_bacteria)))>thresh) == TRUE))
  pc_cluster_bacteria <- pca_bacteria$x[, 1:nr_pc_bacteria]
} else {
  pc_cluster_bacteria <- hs.norm
}

# Evaluate number of robust clusters by means of silhouette index
# We limit the search to 50 clusters
tmp.si <- c()
for(i in 2:50){
  if(i%%10 == 0) cat(date(), paste0("---- at k =  ", i, "/",  nrow(pc_cluster_bacteria), "\n"))
  tmp.si[i] <- pam(pc_cluster_bacteria, k = i)$silinfo$avg.width
}
nr_clusters_bacteria <- which(tmp.si == max(tmp.si, na.rm = TRUE))

# Plot Silhouette index distribution
plot(tmp.si, type = "l", ylab = "Silhouette index", 
     xlab = "Number of clusters")

# Cluster samples and export cluster labels
clusters_bacteria <- pam(pc_cluster_bacteria, k = nr_clusters_bacteria)

# Extract cluster labels
cluster_labels_pam <- data.frame(Sample = names(clusters_bacteria$clustering),
                                      cluster_label = clusters_bacteria$clustering)

# Method 2: the Mclust( ) function in the mclust package selects the optimal model according to BIC for EM initialized by hierarchical clustering for parameterized Gaussian mixture models.
# BIC = mclustBIC(pc_cluster_bacteria$spc, G = c(1:10), modeNames = c("VVI"))
# plot(BIC)
mc_fit <- Mclust(pc_cluster_bacteria, G = c(1:10))

# plot(fit) # plot results 
summary(mc_fit) # display the best model

cluster_labels_mc <- data.frame(Sample = names(clusters_bacteria$clustering),
                                      cluster_label = mc_fit$classification)

# To compare both clustering approaches:
# cluster.stats(dist(hs.norm), mc_fit$classification, clusters_bacteria$clustering)

# Extract count table (i.e. "operational phenotypic unit table") for each sample
OPU_hs_pam <- data.frame(table(cluster_labels_pam))
# print(OPU_hs_pam)

OPU_hs_mc <- data.frame(table(cluster_labels_mc))
# print(OPU_hs_mc)

# Merge cluster outputs in long format df
OPU_hs_merged <- rbind(OPU_hs_pam, OPU_hs_mc)
OPU_hs_merged <- data.frame(OPU_hs_merged, method =
                              c(rep("PAM", nrow(OPU_hs_pam)),
                                rep("Mclust", nrow(OPU_hs_mc))), 
                            replicate = do.call(rbind, strsplit(as.character(OPU_hs_merged$Sample), " "))[, 2],
                            growth_phase = do.call(rbind, strsplit(as.character(OPU_hs_merged$Sample), " "))[, 1])
colnames(OPU_hs_merged)[colnames(OPU_hs_merged) == "cluster_label"] <- "OPU"
```

# Plot OPU table
```{r plot-clusters, fig.width = 7, fig.height= 6, dpi = 500, warning=FALSE}
# Plot according to metadata
p1 <- ggplot(OPU_hs_merged, aes(x = replicate, y = Freq, fill = OPU))+
  geom_bar(stat = "identity")+
  scale_fill_brewer(palette = "Paired")+
  theme_bw()+
  facet_grid(method ~ growth_phase, scales = "free")+
   theme(axis.title=element_text(size=16), strip.text=element_text(size=16),
        legend.title=element_text(size=15),legend.text=element_text(size=14),
        axis.text = element_text(size=14),title=element_text(size=20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background=element_rect(fill=adjustcolor("lightgray",0.2))
        #,panel.grid.major = element_blank(), panel.grid.minor = element_blank()
        )

print(p1)
```

## B. Maldiquant-normalized spectra  

Same analysis as for hyperspec-normalized spectra.

```{r determine-clusters-mq, fig.width = 7, fig.height= 6, dpi = 500, warning=FALSE}
# Convert massSpectrum object to hyperspec
wv_mq <- mass(mq.norm[[1]])
matrix.spectra <- matrix(nrow=length(mq.norm), ncol = length(wv_mq))
for (i in 1:length(mq.norm)){
  matrix.spectra[i,] <- intensity(mq.norm[[i]])
}
hs.mq <- new("hyperSpec", spc = matrix.spectra, wavelength = wv_mq, labels = cell.name)

# Choose if you want to run PCA prior to clustering
PCA <- FALSE
PEAKS <- FALSE

if(PCA == TRUE){
  # Perform PCA to reduce number of features in fingerprint
  pca_bacteria <- prcomp(hs.mq)

  # Only retain PC which explain 90% of the variance
  thresh <- 0.9
  nr_pc_bacteria <- min(which((cumsum(vegan::eigenvals(pca_bacteria)/sum(vegan::eigenvals(pca_bacteria)))>thresh) == TRUE))
  pc_cluster_bacteria <- pca_bacteria$x[, 1:nr_pc_bacteria]
} else if(PEAKS == TRUE){
  # Run peak detection algorithm
    peaks <- detectPeaks(mq.norm, method="MAD", halfWindowSize=1, SNR=0.001)
    plot(mq.norm[[1]], xlim=c(600, 1800))
    points(peaks[[1]], col="red", pch=4)

    # Tolerance for wave number shift
    peaks <- binPeaks(peaks, tolerance = 0.002)
  
    # Filter out intensities at peak wave numbers
    peaks <- filterPeaks(peaks, minFrequency = 0.25)

    pc_cluster_bacteria <- intensityMatrix(peaks, mq.norm)
} else {
  pc_cluster_bacteria <- hs.mq
}

# Evaluate number of robust clusters by means of silhouette index
# We limit the search to 50 clusters
tmp.si <- c()
for(i in 2:50){
  if(i%%10 == 0) cat(date(), paste0("---- at k =  ", i, "/",  nrow(pc_cluster_bacteria), "\n"))
  tmp.si[i] <- pam(pc_cluster_bacteria, k = i)$silinfo$avg.width
}
nr_clusters_bacteria <- which(tmp.si == max(tmp.si, na.rm = TRUE))

# Plot Silhouette index distribution
plot(tmp.si, type = "l", ylab = "Silhouette index", 
     xlab = "Number of clusters")

# Cluster samples and export cluster labels
clusters_bacteria <- pam(pc_cluster_bacteria, k = nr_clusters_bacteria)

# Extract cluster labels
cluster_labels_pam <- data.frame(Sample = cell.name,
                                      cluster_label = clusters_bacteria$clustering)

# Method 2: the Mclust( ) function in the mclust package selects the optimal model according to BIC for EM initialized by hierarchical clustering for parameterized Gaussian mixture models.
if(PEAKS == TRUE){
  mc_fit <- Mclust(as.matrix(pc_cluster_bacteria))
} else {
  mc_fit <- Mclust(pc_cluster_bacteria, G = c(1:10))
}

# plot(fit) # plot results 
summary(mc_fit) # display the best model

cluster_labels_mc <- data.frame(Sample = cell.name,
                                      cluster_label = mc_fit$classification)

# To compare both clustering approaches:
# cluster.stats(dist(hs.mq), mc_fit$classification, clusters_bacteria$clustering)

# Extract count table (i.e. "operational phenotypic unit table") for each sample
OPU_mq_pam <- data.frame(table(cluster_labels_pam))
# print(OPU_hs_pam)

OPU_mq_mc <- data.frame(table(cluster_labels_mc))
# print(OPU_hs_mc)

# Merge cluster outputs in long format df
OPU_mq_merged <- rbind(OPU_mq_pam, OPU_mq_mc)
OPU_mq_merged <- data.frame(OPU_mq_merged, method =
                              c(rep("PAM", nrow(OPU_mq_pam)),
                                rep("Mclust", nrow(OPU_mq_mc))), 
                            replicate = do.call(rbind, strsplit(as.character(OPU_mq_merged$Sample), " "))[, 2],
                            growth_phase = do.call(rbind, strsplit(as.character(OPU_mq_merged$Sample), " "))[, 1])
colnames(OPU_mq_merged)[colnames(OPU_mq_merged) == "cluster_label"] <- "OPU"
```

# Plot OPU table
```{r  plot-clusters-mq, fig.width = 7, fig.height= 6, dpi = 500, warning=FALSE}
p2 <- ggplot(OPU_mq_merged, aes(x = replicate, y = Freq, fill = OPU))+
  geom_bar(stat = "identity")+
  scale_fill_brewer(palette = "Paired")+
  theme_bw()+
  facet_grid(method ~ growth_phase, scales = "free")+
   theme(axis.title=element_text(size=16), strip.text=element_text(size=16),
        legend.title=element_text(size=15),legend.text=element_text(size=14),
        axis.text = element_text(size=14),title=element_text(size=20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background=element_rect(fill=adjustcolor("lightgray",0.2))
        #,panel.grid.major = element_blank(), panel.grid.minor = element_blank()
        )

print(p2)

```

# PhenoRam diversity
```{r  calculate-pheno-div-ram, fig.width = 7, fig.height= 6, dpi = 500, warning=FALSE}
# Format PAM clusters into otu tables
OPU_pam_table <- OPU_mq_merged %>% filter(method == "PAM") %>% select(c("Sample","OPU","Freq")) %>% tidyr::spread(OPU, Freq)
rownames(OPU_pam_table) <- OPU_pam_table$Sample
OPU_pam_table <- OPU_pam_table[, -1]
OPU_pam_tax <- as.matrix(data.frame(OPU = colnames(OPU_pam_table)))
rownames(OPU_pam_tax) <- OPU_pam_tax[,1]
OPU_pam_table2 <- phyloseq(otu_table(OPU_pam_table, taxa_are_rows = FALSE),
                           tax_table(OPU_pam_tax))

# Format Mclust clusters into otu tables
OPU_mclust_table <- OPU_mq_merged %>% filter(method == "Mclust") %>% select(c("Sample","OPU","Freq")) %>% tidyr::spread(OPU, Freq)
rownames(OPU_mclust_table) <- OPU_mclust_table$Sample
OPU_mclust_table <- OPU_mclust_table[, -1]
OPU_mclust_tax <- as.matrix(data.frame(OPU = colnames(OPU_mclust_table)))
rownames(OPU_mclust_tax) <- OPU_mclust_tax[,1]
OPU_mclust_table <- phyloseq(otu_table(OPU_mclust_table, taxa_are_rows = FALSE),
                           tax_table(OPU_mclust_tax))

div_ram_pam <- Diversity_16S(OPU_mclust_table, R = 100, brea = FALSE, 
                             parallel = TRUE, ncores = 3)
div_ram_mclust <- Diversity_16S(OPU_pam_table2, R = 100, brea = FALSE, 
                             parallel = TRUE, ncores = 3)

div_ram_merged <- data.frame(Sample = rep(rownames(div_ram_pam),2),
                             rbind(div_ram_pam, div_ram_mclust),
                             method = c(rep("pam", nrow(div_ram_pam)),
                                        rep("mclust", nrow(div_ram_mclust)))
                             )
div_ram_merged <- div_ram_merged[, -c(4:7)]
div_ram_merged$Sample <- as.character(div_ram_merged$Sample)

# Merge with metadata
div_ram_merged$GrowthPhase <- do.call(rbind, strsplit(div_ram_merged$Sample, " "))[,1]

# Plot results
p_ram_div_pam <- div_ram_merged %>% filter(method == "pam") %>% 
  ggplot(aes(x = GrowthPhase, y = D2, fill = GrowthPhase))+
  geom_point(shape = 21, size = 4)+
  geom_boxplot(alpha = 0.4)+
  ggplot2::theme_bw()+
     theme(axis.title=element_text(size=16), strip.text=element_text(size=16),
        legend.title=element_text(size=15),legend.text=element_text(size=14),
        axis.text = element_text(size=14),title=element_text(size=20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background=element_rect(fill=adjustcolor("lightgray",0.2))
        #,panel.grid.major = element_blank(), panel.grid.minor = element_blank()
        )+
  scale_fill_brewer(palette = "Accent")+
  geom_errorbar(aes(ymin = D2 - sd.D2, ymax = D2 + sd.D2), width = 0.025)+
  guides(fill = FALSE)+
  ylab(expression("Phenotypic diversity - D2 (Raman)"))

print(p_ram_div_pam)

p_ram_div_mclust <- div_ram_merged %>% filter(method == "mclust") %>% 
  ggplot(aes(x = GrowthPhase, y = D2, fill = GrowthPhase))+
  geom_point(shape = 21, size = 4)+
  geom_boxplot(alpha = 0.4)+
  ggplot2::theme_bw()+
     theme(axis.title=element_text(size=16), strip.text=element_text(size=16),
        legend.title=element_text(size=15),legend.text=element_text(size=14),
        axis.text = element_text(size=14),title=element_text(size=20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background=element_rect(fill=adjustcolor("lightgray",0.2))
        #,panel.grid.major = element_blank(), panel.grid.minor = element_blank()
        )+
  scale_fill_brewer(palette = "Accent")+
  geom_errorbar(aes(ymin = D2 - sd.D2, ymax = D2 + sd.D2), width = 0.025)+
  guides(fill = FALSE)+
  ylab(expression("Phenotypic diversity - D2 (Raman)"))

print(p_ram_div_mclust)
```

# Contrast analysis
## Hyperspec normalized  

```{r  plot-contrasts-hs, fig.width = 10, fig.height= 6, dpi = 500, warning=FALSE}
# ram_contrast(hyprs = hs.norm, comp1 = c("LB rep1", "LB rep2", "LB rep3"), 
# comp2 = c("NB rep1","NB rep2","NB rep3"))
ram.hs_lag_log <- ram_contrast(hs.norm, comp1 = c("lag rep1", "lag rep2", "lag rep3"),
              comp2 = c("log rep1", "log rep2", "log rep3"), plot = FALSE)

ram.hs_lag_stat <- ram_contrast(hs.norm, comp1 = c("lag rep1", "lag rep2", "lag rep3"),
              comp2 = c("stat rep1", "stat rep2", "stat rep3"), plot = FALSE)

ram.hs_log_stat <- ram_contrast(hs.norm, comp1 = c("log rep1", "log rep2", "log rep3"),
              comp2 = c("stat rep1", "stat rep2", "stat rep3"), plot = FALSE)

ram.hs_merged <- data.frame(rbind(ram.hs_lag_log, ram.hs_lag_stat, ram.hs_log_stat),
                            Comparison = rep(c("lag-log", "lag-stat", "log-stat"), 
                                               each = nrow(ram.hs_lag_stat))
)

v.hs <- ggplot2::ggplot(ram.hs_merged, ggplot2::aes(x = Wavenumber, y = Density, fill = Density))+
  ggplot2::geom_point(shape = 21, colour="black", alpha = 1.0,
                          size = 3)+
  geom_line()+
  facet_grid(.~Comparison)+
  ggplot2::scale_fill_distiller(palette="RdBu", na.value="white", limits = c(-0.22,0.22)) +
  scale_x_continuous(breaks = seq(600,1800,200), labels = seq(600,1800,200))+
  ggplot2::theme_bw()+
     theme(axis.title=element_text(size=16), strip.text=element_text(size=16),
        legend.title=element_text(size=15),legend.text=element_text(size=14),
        axis.text = element_text(size=14),title=element_text(size=20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background=element_rect(fill=adjustcolor("lightgray",0.2))
        #,panel.grid.major = element_blank(), panel.grid.minor = element_blank()
        )

print(v.hs)
```

## Maldiquant normalized  

```{r  plot-contrasts-mq, fig.width = 10, fig.height= 6, dpi = 500, warning=FALSE}
# ram_contrast(hyprs = hs.norm, comp1 = c("LB rep1", "LB rep2", "LB rep3"), 
# comp2 = c("NB rep1","NB rep2","NB rep3"))
ram.mq_lag_log <- ram_contrast(hs.mq, comp1 = c("lag rep1", "lag rep2", "lag rep3"),
              comp2 = c("log rep1", "log rep2", "log rep3"), plot = FALSE)

ram.mq_lag_stat <- ram_contrast(hs.mq, comp1 = c("lag rep1", "lag rep2", "lag rep3"),
              comp2 = c("stat rep1", "stat rep2", "stat rep3"), plot = FALSE)

ram.mq_log_stat <- ram_contrast(hs.mq, comp1 = c("log rep1", "log rep2", "log rep3"),
              comp2 = c("stat rep1", "stat rep2", "stat rep3"), plot = FALSE)

ram.mq_merged <- data.frame(rbind(ram.mq_lag_log, ram.mq_lag_stat, ram.mq_log_stat),
                            Comparison = rep(c("lag-log", "lag-stat", "log-stat"), 
                                               each = nrow(ram.mq_lag_stat))
)

v.mq <- ggplot2::ggplot(ram.mq_merged, ggplot2::aes(x = Wavenumber, y = Density, fill = Density))+
  ggplot2::geom_point(shape = 21, colour="black", alpha = 1.0,
                          size = 3)+
  geom_line()+
  facet_grid(.~Comparison)+
  ggplot2::scale_fill_distiller(palette="RdBu", na.value="white", limits = c(-0.22,0.22)) +
  scale_x_continuous(breaks = seq(600,1800,200), labels = seq(600,1800,200))+
  ggplot2::theme_bw()+
     theme(axis.title=element_text(size=16), strip.text=element_text(size=16),
        legend.title=element_text(size=15),legend.text=element_text(size=14),
        axis.text = element_text(size=14),title=element_text(size=20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background=element_rect(fill=adjustcolor("lightgray",0.2))
        #,panel.grid.major = element_blank(), panel.grid.minor = element_blank()
        )

print(v.mq)
```

# Flow cytometry data

### Alpha diversity analysis  


```{r  FCM-analysis-1, fig.width = 10, fig.height= 6, dpi = 500, warning=FALSE}
# Import data in FCS format
fs <- flowCore::read.flowSet(path = "./FCSfiles", pattern = ".fcs")

# Extract metadata from sample names
meta_fs <- do.call(rbind, strsplit(flowCore::sampleNames(fs), " "))[,2]
meta_fs <- data.frame(Sample = flowCore::sampleNames(fs),
                      Organism = do.call(rbind, strsplit(meta_fs, "_"))[,1],
                      GrowthPhase = do.call(rbind, strsplit(meta_fs, "_"))[,2],
                      Replicate= do.call(rbind, strsplit(meta_fs, "_"))[,3],
                      Dilution= do.call(rbind, strsplit(meta_fs, "_"))[,4])
meta_fs$Dilution <- as.numeric(gsub(".fcs", "", meta_fs$Dilution))

# Transform data with asinh
# Select phenotypic features of interest and transform parameters
fs <- flowCore::transform(fs,`FL1-H`=asinh(`FL1-H`), 
                                   `SSC-H`=asinh(`SSC-H`), 
                                   `FL3-H`=asinh(`FL3-H`), 
                                   `FSC-H`=asinh(`FSC-H`))
param=c("FL1-H", "FL3-H","SSC-H","FSC-H")

# Denoise data
### Create a PolygonGate for denoising the dataset
### Define coordinates for gate in sqrcut1 in format: c(x,x,x,x,y,y,y,y)
sqrcut1 <- matrix(c(7.75,7.75,14,14,3,7.75,14,3),ncol=2, nrow=4)
colnames(sqrcut1) <- c("FL1-H","FL3-H")
polyGate1 <- polygonGate(.gate=sqrcut1, filterId = "Total Cells")

###  Gating quality check
xyplot(`FL3-H` ~ `FL1-H`, data=fs[1], filter=polyGate1,
       scales=list(y=list(limits=c(0,14)),
                   x=list(limits=c(6,16))),
       axis = axis.default, nbin=125, 
       par.strip.text=list(col="white", font=2, cex=2), smooth=FALSE)

### Isolate only the cellular information based on the polyGate1
fs <- Subset(fs, polyGate1)

#Normalize
summary <- fsApply(x = fs, FUN = function(x) apply(x, 2, max), use.exprs = TRUE)
maxval <- max(summary[,9])
mytrans <- function(x) x/maxval
fs <- transform(fs,`FL1-H`=mytrans(`FL1-H`),
                                  `FL3-H`=mytrans(`FL3-H`), 
                                  `SSC-H`=mytrans(`SSC-H`),
                                  `FSC-H`=mytrans(`FSC-H`))

# Calculate phenotypic diversity
fs_div <- Diversity_rf(fs, param = param, R.b = 100, R = 100, cleanFCS = FALSE,
                       parallel = TRUE, ncores = 3)

## Plot alpha diversity vs growth phase
fs_div <- dplyr::left_join(fs_div, meta_fs, by = c("Sample_names"="Sample"))

p_fs_div <- ggplot(fs_div, aes(x = GrowthPhase, y = D2, fill = GrowthPhase))+
  geom_point(shape = 21, size = 4)+
  geom_boxplot(alpha = 0.4)+
  ggplot2::theme_bw()+
     theme(axis.title=element_text(size=16), strip.text=element_text(size=16),
        legend.title=element_text(size=15),legend.text=element_text(size=14),
        axis.text = element_text(size=14),title=element_text(size=20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background=element_rect(fill=adjustcolor("lightgray",0.2))
        #,panel.grid.major = element_blank(), panel.grid.minor = element_blank()
        )+
  scale_fill_brewer(palette = "Accent")+
  geom_errorbar(aes(ymin = D2 - sd.D2, ymax = D2 + sd.D2), width = 0.025)+
  guides(fill = FALSE)+
  ylab(expression("Phenotypic diversity - D2 (FCM)"))

# Compare FCM diversity (left hand side) with Raman diversity (right hand side)
grid.arrange(p_fs_div, p_ram_div_mclust, ncol = 2)
```

### Contrast analysis  

```{r  FCM-analysis-2, fig.width = 10, fig.height= 6, dpi = 500, warning=FALSE}
# Calculate fingerprint
fbasis <- flowBasis(fs, param, nbin=128, 
                   bw=0.01,normalize=function(x) x)

# Calculate contrasts
fp_c_lag_log <- fp_contrasts(fbasis, comp1 = meta_fs$GrowthPhase=="lag", comp2 = 
               meta_fs$GrowthPhase=="log", thresh = 0.05)
fp_c_lag_stat <- fp_contrasts(fbasis, comp1 = meta_fs$GrowthPhase=="lag", comp2 = 
               meta_fs$GrowthPhase=="stat", thresh = 0.05)
fp_c_log_stat <- fp_contrasts(fbasis, comp1 = meta_fs$GrowthPhase=="log", comp2 = 
               meta_fs$GrowthPhase=="stat", thresh = 0.05)

fp_c_merge <- data.frame(rbind(fp_c_lag_log, fp_c_lag_stat, fp_c_log_stat),
                         Comparison = c(rep("lag-log", nrow(fp_c_lag_log)),
                                        rep("lag-stat", nrow(fp_c_lag_stat)),
                                        rep("log-stat", nrow(fp_c_log_stat))))
# Plot contrasts
v.fp_c <- ggplot2::ggplot(fp_c_merge, ggplot2::aes(x = FL1.H, y = FL3.H, fill = Density, 
                                                   z = Density))+
  geom_tile(aes(fill=Density)) + 
  geom_point(colour="gray", alpha=0.7)+
  scale_fill_distiller(palette="RdYlBu", na.value="white") + 
  stat_contour(aes(fill=..level..), geom="polygon", binwidth=0.1)+
  theme_bw()+
  # geom_line()+
  facet_grid(.~Comparison)+
  # scale_x_continuous(breaks = seq(600,1800,200), labels = seq(600,1800,200))+
  theme(axis.title=element_text(size=16), strip.text=element_text(size=16),
        legend.title=element_text(size=15),legend.text=element_text(size=14),
        axis.text = element_text(size=14),title=element_text(size=20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background=element_rect(fill=adjustcolor("lightgray",0.2))
        #,panel.grid.major = element_blank(), panel.grid.minor = element_blank()
        )+
  labs(x="Green fluorescence intensity (a.u.)", y="Red fluorescence intensity (a.u.)")

print(v.fp_c)
```

### Beta diversity analysis

```{r  FCM-analysis-3, fig.width = 7, fig.height= 6, dpi = 500, warning=FALSE}
# Calculate beta diversity
beta.div <- beta_div_fcm(fbasis, ord.type="PCoA")

# Plot ordination
plot_beta_fcm(beta.div, color = meta_fs$GrowthPhase, labels="Growth phase") + 
  theme_bw() +
  geom_point(size = 8, alpha = 0.5)+
  ggtitle("")

# Juxtaposition with beta-diversity based on Raman data
```